### 2021-11-25 review/response-RC2.tex
Add remark about do3se model description.

% Authors' response
\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{multirow}
\thispagestyle{empty}

\begin{document}
\section*{Authors' response}
To bg-2021-260-RC2 (23 Nov 2021):
We thank the anonymous referee \#2 for their very constructive comments and useful suggestions.
We will address the \textbf{major comments} raised regarding the structure and readability of the manuscript in our next revision. We will refine the language to reduce ambiguity and supplement our statements with appropriate citations.\\
As the manuscript will, hence, undergo major structural changes, we have to refrain from addressing all \textbf{specific comments} in detail at this point. In the final revision of the manuscript, we will take all specific and technical comments which are still relevant into consideration.\\
In the following, we shall briefly respond to all issues where it seems appropriate.
\subsection*{General comments}
\begin{itemize}

    \item {\color{blue} "[...] main figure is missing [...]"} The referee is referring to Fig.~8 which is apparently absent from their version of the manuscript. We downloaded the pre-print from the associated discussion page (\href{https://bg.copernicus.org/preprints/bg-2021-260/}{bg-2021-260}) and cannot confirm Fig.~8 to be missing. It can be found on page~23 (just before the list of references). We, however, acknowledge that the current placement (after the main body of text) is rather unfortunate. This shall be fixed in the final typesetting process!

\end{itemize}

\subsection*{Specific comments}
\begin{itemize}

    \item {\color{blue} "[...] I find the introduction to be a bit longer than would be most effective for conveying this information. It would be helpful to break it into subsections [...]. [...]content of [...] lines 115--134 is important for contextualizing [...] could be better emphasized [...] or [...] opening paragraph for [...] section~2 [...]."} Thank you for this constructive comment. We will condense our introduction and put subsections in place where necessary. As suggested, we move the discussion of meteorology to Section~2.
    \item {\color{blue} "Please correct the text in line 105. [...] This doesn't seem to be a correct descriptor for Equation~2 [...]"} Thank you for pointing out this carelessness which should not have gone unnoticed.
    \item {\color{blue} "Section~4 - Please expand the description of the $\mathrm{DO_3SE}$ model"} A comprehensive model description can be found in Appendix~B. As no modifications to the $\mathrm{DO_3SE}$ model were made (except for the change in parameters), we had concluded that an excessive model description would distract the reader. In particular, relationships between plant physiological parameters and meteorological observations are described in Appendix~B1. Model "high-level" input and its preparation is described in Appendix~B2. We agree that important information is therefore missing from the main text. In the course of restructuring, we include a condensed model description in Section~4.
    \item {\color{blue} "$f_\mathrm{T}$ hasn't been defined yet." "Is $f_\mathrm{T} = f_\mathrm{temp}$?"} These two issues are related. Of course, $f_\mathrm{T}$ is the same as $f_\mathrm{temp}$ and we should not have used both at the same time. We remove this ambiguity.
    \item {\color{blue} "Line 317-318 -- Mention that this is PPFD~0.8 ans 1.2 in parentheses [...]"} We add the suggested information to clarify our naming convention.
    \item {\color{blue} "Line 321 appears to substitute for $f_\mathrm{light}$" in Equation 4, not for as described in the text."} Indeed this sentence is ambiguous, and should read: "First, we solve Eq.~(3) for the MM default values of $\alpha_\mathrm{MM}$ at $f_\mathrm{light}=0.5$" We should, perhaps, clarify that we use $\gamma$ synonymously with PPFD in this paragraph or substitute $\gamma$ with PPFD.
    \item {\color{blue} "Line 331 -- Is the low standard deviation specific to nighttime?"} No, we have selected our sample (May--October), so that the vegetation will not experience "nighttime" between 5~am and 9~am. See, f.e. \href{https://www.timeanddate.com/sun/@777232}{TimeAndDate.com/Svanvik}.
    \item {\color{blue} "I find the conclusion that the standard deviation indicating "higher robustness to variability in growing conditions" to sweeping relative to the evidence precented. [...]"} This might be a misunderstanding due to inaccuracies in language? The robustness is not a conclusion but a presumption. We assume that the essential climate variables (excluding ozone) vary substantially in our sampled 20 years. Though, we have not explicitly shown the associated standard deviation of this subsample of our data. We could do this to strengthen our point. A low standard deviation in stomatal conductance, therefore, suggests that stomatal conductance is relatively unaffected by these variations. We refer to this as "robustness". We will make this more clear in the next revision. Or remove the whole point as suggested.
    \item {\color{blue} "Line 332 -- "subarctic-PPFD0.8" is "best" relative to what?"} In this case "best" refers to the highest relative stomatal conductance and smallest standard deviation compared to the other probed combinations of parameter. We rephrase the sentence accordingly.
    \item {\color{blue}"Figure~8 appears to be missing from the document."} As mentioned earlier, the placement of the figure on page 23 (after the main body of text) is unfortunate. The figure is not missing (at least as far as we can see), but might, therefore, have slipped the vigilant eye of the referee.

\end{itemize}

\subsection*{Technical comments}
We will appropriately address all comments raised, but only address selected items at this point.
\begin{itemize}
    \item {\color{blue} Concerning Figures 3--5} Thank you for the useful suggestions to further improve our figures. We will take them into consideration. We may kindly ask to review Figure~8 as well.
    \item {\color{blue} I haven't previously seen this method for identifying the start and end points of photosynthesis, and (noting that I'm not an ecologist) it seems elegant in concept and approach.} Thank you. This was a pragmatic, data-driven approach. To our knowledge, most satellite data based studies on start and end of growing season focus on correlations with essential climate variables to identify processes and relationships.
    \item {\color{blue} "[...] please clarify what leaf length you are using [...]"} We have conducted two sets of simulations, one with leaf dimensions, tree height, and growing season as defined in the mapping manual and one with leaf dimensions of deciduous trees adjusted to the measured size. We have not adjusted the leaf dimensions of coniferous trees and grassland. We will make this clearer.
    \item {\color{blue} "Line 457 -- do you mean the response to SWP, rather SWP itself being negligible?"} Of course! We apologize for the sloppy language.

\end{itemize}

\end{document}