### 2021-12-20 review/response-RC3.tex
Init response to review 3

% Authors' response
\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{multirow}
\thispagestyle{empty}

\begin{document}
\section*{Authors' response}
To bg-2021-260-RC3 (30 Nov 2021):
We thank the anonymous referee \#3 for their comments.
We will address all \textbf{general comments} raised regarding the structure and readability of the manuscript in our next revision or resubmission. We consider to separate the manuscript in Introduction--Methods--Results--Conclusions and will improve on the "methods" Section to make our results more comprehensible. We will refine the language to reduce ambiguity and supplement our statements with appropriate citations where indicated by the referee.\\
As the manuscript will, hence, undergo major structural changes, we will not address all \textbf{specific / technical comments} in detail at this point. We will take all specific and technical comments which are still relevant into consideration.\\
In the following, we shall give a brief respond to all relevant issues.

\subsection*{Major comments} 
\begin{itemize}

    \item {\color{blue}Concerning the length of the paper, I really cannot see the benefit of all the work on long-
    term ozone and meteorology given in Sections 2-3. Isn't the main point of this paper to
    illustrate differences in vegetation characteristics between plants at nearly 70 degrees N
    and the Mapping Manual (MM)? That doesn't need a long analysis of climate differences or
    ozone trends.}

    \item {\color{blue}Concerning the "default" growing seasons (from DOY 100 to 307) from the MM against
    which the "bespoke" are compared, these are just wrong for this latitude. The MM
    suggests a latitude function that gives DOY 129 to 268.}

    \item {\color{blue}Some of the basic equations look wrong to me, see my comments on Appendix B below.}

    \item {\color{blue}On p21 the author's first "key findings" state that their "bespoke parameterizations for
    subarctic species are better suited..." than the MM ones. This is a strong statement given
    that the authors admit on L293 that their bespoke suggestions are "hypothetical and have
    yet to be verified by observations".}

\end{itemize}

\subsection*{Other comments} 
\begin{itemize}

\item {\color{blue}b}

\end{itemize}

\end{document}